---
title: "LTDE BGE Experiment"
author: "Mario E. Muscarella, Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
   - \usepackage{graphics}
output: 
  pdf_document:
    fig_caption: true
geometry: margin=2.54cm
---

```{r, echo=FALSE, results='hide'}
rm(list=ls())
setwd("~/GitHub/StarvationTraits/analyses/")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
CV <- function(x, ...){(sd(x, na.rm = TRUE)/mean(x, na.rm = TRUE))*100}
library("png")
library("grid")
library("reshape")
```

# Overview

Is growth efficiency an evolvable trait? Here, we explore patterns in bacterial 
growth efficiency using a set of isolates from the long-term starvation experiment. 
Each isolate has an ancestor and at least three derived strains that were isolated
after three years of starvation in PBS.
We are using 14C succinate as the sole carbon source for these experiments.

# Experiment 1: BGE Evolution Experiment

## Data Import 
```{r, results='hide'}
# Import Data
# KBS0707.data <- read.csv("../data/BGE/20160216_BRBP.csv", header = T)
# Not using this run because it doesn't appear that the cultures grew enough
KBS0707.data <- read.csv("../data/BGE/20160228_BRBP.csv", header = T)
KBS0802.data <- read.csv("../data/BGE/20160214_BRBP.csv", header = T)

```

## Data Processing
```{r, results='hide'}
# Raw Data For Intra Vial Test
voucher707  <- KBS0707.data[grep("Voucher", KBS0707.data$Tube), ]
voucher802  <- KBS0802.data[grep("Voucher", KBS0802.data$Tube), ]
control707   <- KBS0707.data[grep("Control", KBS0707.data$Tube), ]
control802   <- KBS0802.data[grep("Control", KBS0802.data$Tube), ]

# Merge Datasets
BGE.data.full <- rbind(KBS0707.data, KBS0802.data)

# Remove Control and Voucher
BGE.data <- BGE.data.full[-which(BGE.data.full$Tube == "Voucher" | 
                                   BGE.data.full$Tube == "Control"), ]
BGE.data$CPM.c <- NA

for (i in 1:dim(BGE.data)[1]){
  if (BGE.data$Sample[i] == "Cells"){
    BGE.data$CPM.c[i] <- BGE.data$CPM[i] * (6150/1500)
  } else {
  if (BGE.data$Sample[i] == "Trap"){
    BGE.data$CPM.c[i] <- BGE.data$CPM[i] * (500/100)
  } else {
  if (BGE.data$Sample[i] == "Sup"){
    BGE.data$CPM.c[i] <- BGE.data$CPM[i] * (6150/100)
  }
  else {
    BGE.data$CPM.c[i] <- NA
  }}}
}

BGE.data.m <- melt(BGE.data)
BGE.data.p <- cast(BGE.data.m, Tube + Organism + Sample ~ variable, 
                   subset=variable=="CPM.c",  fun.aggregate = c(mean, CV))

BGE.data.p$CPM.p <- NA
for (i in 1:dim(BGE.data.p)[1]){
  if (grepl("707", BGE.data.p$Organism[i])){
    BGE.data.p$CPM.p <- BGE.data.p$CPM.c_mean / voucher707$CPM
  } else {
  if (grepl("802", BGE.data.p$Organism[i])){
    BGE.data.p$CPM.p <- BGE.data.p$CPM.c_mean / voucher802$CPM
  } else{
    BGE.data.p$CPM.p <- NA
  }}} 

mass.l <- melt(as.data.frame(BGE.data.p))
# Mass Balance
mass <- cast(mass.l, Tube + Organism ~ variable, subset=variable == "CPM.p", 
             fun.aggregate = sum)
mean(mass$CPM.p)

# Data For Plots
BGE.707 <- droplevels(BGE.data.p[grep("707", BGE.data.p$Organism), ])
BGE.802 <- droplevels(BGE.data.p[grep("802", BGE.data.p$Organism), ])
levels(BGE.802$Tube)
```

## 802 Plot
```{r}
BGE.802.w <- reshape(BGE.802, idvar = "Tube", direction="wide", 
                     v.names = "CPM.p", timevar = "Sample", 
                     drop = c("CPM.c_mean", "CPM.c_CV"))

BGE.802.w$BGE <- BGE.802.w$CPM.p.Cells / (BGE.802.w$CPM.p.Cells + 
                                            BGE.802.w$CPM.p.Trap)
BGE.802.m <- melt(data.frame(BGE.802.w[,c(1,2,6)]))
BGE.802.c <- cast(BGE.802.m, Organism ~ variable, fun.aggregate=c(mean, se))

# Plot
png(filename="../figures/BGE802.png",
    width = 1000, height = 800, res = 96*2)

par(mar=c(3,5.5,2.5,1))
bge_plot <- plot(BGE.802.c$BGE_mean, main = "LTDE: KBS0802 BGE", 
                 pch = 15, cex = 1.5,
        ylab = "", yaxt= "n", xaxt = "n", xlab = "",
        xlim = c(0.5, 4.5), ylim = c(0.26, 0.33))

arrows(x0 = c(1:4), y0 = BGE.802.c$BGE_mean, y1 = BGE.802.c$BGE_mean - 
         BGE.802.c$BGE_se, angle = 90,
       length=0.1, lwd = 2)
arrows(x0 = c(1:4), y0 = BGE.802.c$BGE_mean, y1 = BGE.802.c$BGE_mean + 
         BGE.802.c$BGE_se, angle = 90,
       length=0.1, lwd = 2)

axis(side = 1, labels = c("Ancestor", "A-1", "B-1", "C-1"), 
     at = c(1:4), lwd.ticks = 2, cex = 2)
axis(side = 2, labels = T, at = c(0.26, 0.28, 0.30, 0.32), las = 1, 
     lwd.ticks = 2)
axis(side=1, lwd.ticks = 2, tck = -0.02, labels=F, cex.axis=1, at = c(1:4))
axis(side=3, lwd.ticks = 2, tck = -0.02, labels=F, cex.axis=1, at = c(1:4))
axis(side=1, lwd.ticks = 2, tck = 0.01, labels=F, cex.axis=1, at = c(1:4))
axis(side=3, lwd.ticks = 2, tck = 0.01, labels=F, cex.axis=1, at = c(1:4))
axis(side = 2, lwd.ticks = 2, tck = -0.02, labels=F, cex.axis=1, 
     at = c(0.26, 0.28, 0.30, 0.32))
axis(side = 4, lwd.ticks = 2, tck = -0.02, labels=F, cex.axis=1, 
     at = c(0.26, 0.28, 0.30, 0.32))
axis(side = 2, lwd.ticks = 2, tck = 0.01, labels=F, cex.axis=1, 
     at = c(0.26, 0.28, 0.30, 0.32))
axis(side = 4, lwd.ticks = 2, tck = 0.01, labels=F, cex.axis=1, 
     at = c(0.26, 0.28, 0.30, 0.32))

box(lwd = 2)

mtext("BGE", side = 2 , line =3.5, cex = 1.5)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="14C-Succinate Recovery and BGE"}
img <- readPNG("../figures/BGE802.png")
 grid.raster(img)
```

## 707 Plot
```{r}
BGE.707.w <- reshape(BGE.707, idvar = "Tube", direction="wide", 
                     v.names = "CPM.p", timevar = "Sample", 
                     drop = c("CPM.c_mean", "CPM.c_CV"))

BGE.707.w$BGE <- BGE.707.w$CPM.p.Cells / (BGE.707.w$CPM.p.Cells + 
                                            BGE.707.w$CPM.p.Trap)
BGE.707.m <- melt(data.frame(BGE.707.w[,c(1,2,6)]))
BGE.707.c <- cast(BGE.707.m, Organism ~ variable, 
                  fun.aggregate=c(mean, se))

# Plot
png(filename="../figures/BGE707.png",
    width = 1000, height = 800, res = 96*2)

par(mar=c(3,5.5,2.5,1))
bge_plot <- plot(BGE.707.c$BGE_mean, main = "LTDE: KBS0707 BGE", 
                 pch = 15, cex = 1.5, ylab = "", yaxt= "n", 
                 xaxt = "n", xlab = "",
                 xlim = c(0.5, 4.5), ylim = c(0.22, 0.28))

arrows(x0 = c(1:4), y0 = BGE.707.c$BGE_mean, y1 = BGE.707.c$BGE_mean - 
         BGE.707.c$BGE_se, angle = 90,
       length=0.1, lwd = 2)
arrows(x0 = c(1:4), y0 = BGE.707.c$BGE_mean, y1 = BGE.707.c$BGE_mean + 
         BGE.707.c$BGE_se, angle = 90,
       length=0.1, lwd = 2)

axis(side = 1, labels = c("Ancestor", "A-1", "B-1", "C-1"), 
     at = c(1:4), lwd.ticks = 2, cex = 2)
axis(side = 2, labels = T, at = c(0.22, 0.24, 0.26, 0.28), 
     las = 1, lwd.ticks = 2)
axis(side=1, lwd.ticks = 2, tck = -0.02, labels=F, cex.axis=1, at = c(1:4))
axis(side=3, lwd.ticks = 2, tck = -0.02, labels=F, cex.axis=1, at = c(1:4))
axis(side=1, lwd.ticks = 2, tck = 0.01, labels=F, cex.axis=1, at = c(1:4))
axis(side=3, lwd.ticks = 2, tck = 0.01, labels=F, cex.axis=1, at = c(1:4))
axis(side = 2, lwd.ticks = 2, tck = -0.02, labels=F, cex.axis=1, 
     at = c(0.22, 0.24, 0.26, 0.28))
axis(side = 4, lwd.ticks = 2, tck = -0.02, labels=F, cex.axis=1, 
     at = c(0.22, 0.24, 0.26, 0.28))
axis(side = 2, lwd.ticks = 2, tck = 0.01, labels=F, cex.axis=1, 
     at = c(0.22, 0.24, 0.26, 0.28))
axis(side = 4, lwd.ticks = 2, tck = 0.01, labels=F, cex.axis=1, 
     at = c(0.22, 0.24, 0.26, 0.28))
box(lwd = 2)

mtext("BGE", side = 2 , line =3.5, cex = 1.5)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="14C-Succinate Recovery and BGE"}
img <- readPNG("../figures/BGE707.png")
 grid.raster(img)
```

# Experiment 2: 14C Succinate BGE Using Michaelis-Menten Experiment
## Data Import
```{r, results='hide'}
KBS0802.data <- read.csv("../data/BGE/20160306_BRBP.csv", header = T)
KBS0707.data <- read.csv("../data/BGE/20160430_BRBP.csv", header = T)
KBS0707A1.data <- read.csv("../data/BGE/20160525_BRBP.csv", header = T)
```

## Data Processing
```{r, results='hide'}
# Raw Data For Intra Vial Test
voucher802  <- KBS0802.data[grep("Voucher", KBS0802.data$Tube), ]
control802   <- KBS0802.data[grep("Control", KBS0802.data$Tube), ]
blank802  <- KBS0802.data[grep("Blank", KBS0802.data$Organism), ]
kill802  <- KBS0802.data[grep("KC", KBS0802.data$Organism), ]

# 707
voucher707  <- KBS0707.data[grep("Voucher", KBS0707.data$Tube), ]
control707   <- KBS0707.data[grep("Control", KBS0707.data$Tube), ]
blank707  <- KBS0707.data[grep("Blank", KBS0707.data$Organism), ]
kill707  <- KBS0707.data[grep("KC", KBS0707.data$Organism), ]

voucher707A1  <- KBS0707A1.data[grep("Voucher", KBS0707A1.data$Tube), ]
control707A1   <- KBS0707A1.data[grep("Control", KBS0707A1.data$Tube), ]
blank707A1  <- KBS0707A1.data[grep("Blank", KBS0707A1.data$Organism), ]
kill707A1  <- KBS0707A1.data[grep("KC", KBS0707A1.data$Organism), ]

# Merge Datasets
BGE.data.full <- rbind(KBS0802.data, KBS0707.data, KBS0707A1.data)

# Remove Control and Voucher
BGE.data <- BGE.data.full[-which(BGE.data.full$Tube == "Voucher" | 
                                   BGE.data.full$Tube == "Control" |
                                   BGE.data.full$Organism == "Blank" |
                                   BGE.data.full$Organism == "KC"), ]
BGE.data$CPM.c <- NA

# Turn Conc into Factor
BGE.data$Conc <- as.factor(BGE.data$Conc)

# Create Vector of Final Vol
BGE.data$Vol <- NA
for (i in 1:dim(BGE.data)[1]){
  if (BGE.data$Organism[i] == "KBS0802"){
    BGE.data$Vol[i] = 6150
  } else {
    if (BGE.data$Organism[i] == "KBS0707") {
      BGE.data$Vol[i] = 6200
    } else {
      if (BGE.data$Organism[i] == "KBS0707_A1") {
      BGE.data$Vol[i] = 6200  
    }}
  }
}

# Correct CPM data for entire contents of fraction
for (i in 1:dim(BGE.data)[1]){
  if (BGE.data$Sample[i] == "Cells"){
    BGE.data$CPM.c[i] <- BGE.data$CPM[i] * (BGE.data$Vol[i]/1500)
  } else {
  if (BGE.data$Sample[i] == "Trap"){
    BGE.data$CPM.c[i] <- BGE.data$CPM[i] * (500/100)
  } else {
  if (BGE.data$Sample[i] == "Sup"){
    BGE.data$CPM.c[i] <- BGE.data$CPM[i] * (BGE.data$Vol[i]/100)
  }
  else {
    BGE.data$CPM.c[i] <- NA
  }}}
}

# Re-organize data
BGE.data.m <- melt(BGE.data)
BGE.data.p <- cast(BGE.data.m, Tube + Organism + Conc + 
                     Sample ~ variable, subset=variable=="CPM.c",  
                   fun.aggregate = mean)

# Calculate Percentages given Voucher
BGE.data.p$CPM.p <- NA
for (i in 1:dim(BGE.data.p)[1]){
  if (BGE.data.p$Organism[i] == "KBS0802"){
    BGE.data.p$CPM.p[i] <- BGE.data.p$CPM.c[i] / voucher802$CPM
  } else {
    if (BGE.data.p$Organism[i] == "KBS0707"){
      BGE.data.p$CPM.p[i] <- BGE.data.p$CPM.c[i] / voucher707$CPM
    } else {
      if (BGE.data.p$Organism[i] == "KBS0707_A1"){
      BGE.data.p$CPM.p[i] <- BGE.data.p$CPM.c[i] / voucher707A1$CPM 
    }}
  } 
  }

# Mass Balance
mass.l <- melt(as.data.frame(BGE.data.p))
mass <- cast(mass.l, Tube + Organism + Conc ~ variable, 
             subset=variable == "CPM.p", fun.aggregate = sum)
mean(mass$CPM.p)

# Drop Levels (needed?)
BGE.802 <- droplevels(BGE.data.p[grep("802", BGE.data.p$Organism), ])
BGE.707 <- droplevels(BGE.data.p[grep("707", BGE.data.p$Organism), ])

# Turn percentages into concentrations given experimental concentration
BGE.802.w <- reshape(BGE.802, idvar = "Tube", direction="wide", 
                     v.names = "CPM.p", timevar = "Sample", 
                     drop = c("CPM.c"))
BGE.802.w$CPM.p.Resp <- 1 - (BGE.802.w$CPM.p.Cells + BGE.802.w$CPM.p.Sup)
BGE.802.w$Conc <- as.numeric(as.character(BGE.802.w$Conc))
BGE.802.w$Conc.Cells <- BGE.802.w$CPM.p.Cells * BGE.802.w$Conc
BGE.802.w$Conc.Sup <- BGE.802.w$CPM.p.Sup * BGE.802.w$Conc
BGE.802.w$Conc.Trap <- BGE.802.w$CPM.p.Trap * BGE.802.w$Conc
BGE.802.w$Conc.Resp <- BGE.802.w$CPM.p.Resp * BGE.802.w$Conc

BGE.707.w <- reshape(BGE.707, idvar = c("Organism", "Tube"), direction="wide", 
                     v.names = "CPM.p", timevar = "Sample", 
                     drop = c("CPM.c")) 
BGE.707.w$CPM.p.Resp <- 1 - (BGE.707.w$CPM.p.Cells + BGE.707.w$CPM.p.Sup)
BGE.707.w$Conc <- as.numeric(as.character(BGE.707.w$Conc))
BGE.707.w$Conc.Cells <- BGE.707.w$CPM.p.Cells * BGE.707.w$Conc
BGE.707.w$Conc.Sup <- BGE.707.w$CPM.p.Sup * BGE.707.w$Conc
BGE.707.w$Conc.Trap <- BGE.707.w$CPM.p.Trap * BGE.707.w$Conc
BGE.707.w$Conc.Resp <- BGE.707.w$CPM.p.Resp * BGE.707.w$Conc

# Calculate BGE
BGE.802.w$BGE <- BGE.802.w$CPM.p.Cells / (BGE.802.w$CPM.p.Cells + 
                                            BGE.802.w$CPM.p.Resp)
BGE.707.w$BGE <- BGE.707.w$CPM.p.Cells / (BGE.707.w$CPM.p.Cells + 
                                            BGE.707.w$CPM.p.Resp)

# Recovery
BGE.802.w$Recovery <- rowSums(BGE.802.w[,4:6])
BGE.707.w$Recovery <- rowSums(BGE.707.w[,4:6])
```

## Michaelis Menten Plot KBS802
```{r, results = 'hide'}
# Plot
png(filename="../figures/MM802-14C.png",
    width = 1000, height = 1600, res = 96*2)

par(mfrow = c(1,1), mar = c(0, 6, 0, 0) + 0.5, oma = c(5, 0, 4, 1) + 0.5)
layout(rbind(1, 2, 3), height = c(3, 3, 3)) 

prod_plot <- plot(Conc.Cells ~ jitter(Conc), data = BGE.802.w, 
                 pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5,
                 xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
                 ylim = c(0.0, 0.15))
axis(side=2, lwd.ticks = 2, tck=-0.04, labels=T, las = 1, cex.axis = 1.5)
axis(side=2, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck=-0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.04,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
mtext("BP", side = 2 , line = 4, cex = 1.5)
mtext("KBS0802", side = 3, line = 1.5, outer = F, cex = 1.5)
# mtext("(mM Carbon)", side = 2, line = 3, cex = 0.8)
box(lwd = 3)
resp_plot <- plot(Conc.Resp ~ jitter(Conc), data = BGE.802.w, 
                 pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5,
                 xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
                 ylim = c(0.1, 0.9))
axis(side=2, lwd.ticks = 2, tck=-0.04, labels=T, las = 1, cex.axis = 1.5)
axis(side=2, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck=-0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.04,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
mtext("BR", side = 2 , line = 4, cex = 1.5)
# mtext("(mM Carbon)", side = 2, line = 3, cex = 0.8)
box(lwd = 3)
bge_plot <- plot(BGE ~ Conc, data = BGE.802.w, 
                 pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5,
                 xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
                 ylim = c(0.05, 0.22))
axis(side=2, lwd.ticks = 2, tck=-0.04, labels=T, las = 1, cex.axis = 1.5)
axis(side=2, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck=-0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1.5, las = 1, tck = -0.04,
     at = c(0, 2.5, 5, 7.5, 10), labels = T)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.02,
     at = c(0, 2.5, 5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
mtext("BGE", side = 2 , line = 4, cex = 1.5)
box(lwd = 3)
mtext("Succinate (mM)", side = 1, line = 3.5, outer=F, cex = 1.5)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
#```

#```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="BGE Under Michaelis Menten Dynamics"}
img <- readPNG("../figures/MM802-14C.png")
 grid.raster(img)
```

## Michaelis Menten Plot KBS707
```{r, results = 'hide'}
# Plot
png(filename="../figures/MM707-14C.png",
    width = 1000, height = 1600, res = 96*2)

par(mfrow = c(1,1), mar = c(0, 6, 0, 0) + 0.5, oma = c(5, 0, 4, 1) + 0.5)
layout(rbind(1, 2, 3), height = c(3, 3, 3)) 

prod_plot <- plot(Conc.Cells ~ jitter(Conc), data = BGE.707.w, type = "n",
                 pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5,
                 xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
                 ylim = c(0.01, 0.09), xlim = c(0, 10))
points(Conc.Cells ~ jitter(Conc), data = BGE.707.w[BGE.707.w$Organism == "KBS0707", ],
       pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5)
lines(lowess(BGE.707.w[BGE.707.w$Organism == "KBS0707", ]$Conc, 
             BGE.707.w[BGE.707.w$Organism == "KBS0707", ]$Conc.Cells),
             lty = 2, lwd = 3)
points(Conc.Cells ~ jitter(Conc), data = BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ],
       pch = 21, cex = 3, col = "black", bg = "gray", lwd = 2.5)
lines(lowess(BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ]$Conc, 
             BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ]$Conc.Cells),
             lty = 2, lwd = 3)
axis(side=2, lwd.ticks = 2, tck=-0.04, labels=T, las = 1, cex.axis = 1.5)
axis(side=2, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck=-0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
mtext("BP", side = 2 , line = 4, cex = 1.5)
mtext("KBS0707", side = 3, line = 1.5, outer = F, cex = 1.5)
legend("bottomright", c("Ancestor", "Derived"), pch = c(22, 21), bty = "n",
       pt.bg = "gray", pt.lwd = 2, cex = 1.25, pt.cex = 2, inset = 0.01)
# mtext("(mM Carbon)", side = 2, line = 3, cex = 0.8)
box(lwd = 3)
resp_plot <- plot(Conc.Resp ~ jitter(Conc), data = BGE.707.w, type = "n", 
                 pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5,
                 xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
                 ylim = c(0, 1.5), xlim = c(0, 10))
points(Conc.Resp ~ jitter(Conc), data = BGE.707.w[BGE.707.w$Organism == "KBS0707", ],
       pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5)
lines(lowess(BGE.707.w[BGE.707.w$Organism == "KBS0707", ]$Conc, 
             BGE.707.w[BGE.707.w$Organism == "KBS0707", ]$Conc.Resp),
             lty = 2, lwd = 3)
points(Conc.Resp ~ jitter(Conc), data = BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ],
       pch = 21, cex = 3, col = "black", bg = "gray", lwd = 2.5)
lines(lowess(BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ]$Conc, 
             BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ]$Conc.Resp),
             lty = 2, lwd = 3)
axis(side=2, lwd.ticks = 2, tck=-0.04, labels=T, las = 1, cex.axis = 1.5)
axis(side=2, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck=-0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
mtext("BR", side = 2 , line = 4, cex = 1.5)
# mtext("(mM Carbon)", side = 2, line = 3, cex = 0.8)
box(lwd = 3)
bge_plot <- plot(BGE ~ jitter(Conc), data = BGE.707.w, type = "n",
                 pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5,
                 xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
                 ylim = c(0.02, 0.20), xlim = c(0, 10))
points(BGE ~ jitter(Conc), data = BGE.707.w[BGE.707.w$Organism == "KBS0707", ],
       pch = 22, cex = 3, col = "black", bg = "gray", lwd = 2.5)
lines(lowess(BGE.707.w[BGE.707.w$Organism == "KBS0707", ]$Conc, 
             BGE.707.w[BGE.707.w$Organism == "KBS0707", ]$BGE),
             lty = 2, lwd = 3)
points(BGE ~ jitter(Conc), data = BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ],
       pch = 21, cex = 3, col = "black", bg = "gray", lwd = 2.5)
lines(lowess(BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ]$Conc, 
             BGE.707.w[BGE.707.w$Organism == "KBS0707_A1", ]$BGE),
             lty = 2, lwd = 3)
axis(side=2, lwd.ticks = 2, tck=-0.04, labels=T, las = 1, cex.axis = 1.5)
axis(side=2, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck=-0.02, labels=F)
axis(side=4, lwd.ticks = 2, tck= 0.02, labels=F)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1.5, las = 1, tck = -0.04,
     at = c(0, 2.5, 5, 7.5, 10), labels = T)
axis(side = 1, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = -0.02,
     at = c(0, 2.5, 5, 10), labels = F)
axis(side = 3, lwd.ticks = 2 , cex.axis = 1, las = 1, tck = 0.02,
     at = c(0, 2.5, 5, 7.5, 10), labels = F)
mtext("BGE", side = 2 , line = 4, cex = 1.5)
box(lwd = 3)
mtext("Succinate (mM)", side = 1, line = 3.5, outer=F, cex = 1.5)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="BGE Under Michaelis Menten Dynamics"}
img <- readPNG("../figures/MM707-14C.png")
 grid.raster(img)
```