---
title: "14C Respiration/Production Methods"
author: "Mario Muscarella"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
   - \usepackage{graphics}
output: 
  pdf_document:
    fig_caption: true
geometry: margin=2.54cm
---

```{r, echo=FALSE, results='hide'}
rm(list=ls())
setwd("~/GitHub/StarvationTraits/notes/")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```

# Overview

The simultaneous measurement of bacterial respiration and production will give the best results for estimating growth efficiency.
Here, we have constructed custom bacterial respiration/production vials (Fig 1).
We will use these vials with 14C-resources.
However, first we must optimize the methods.

```{r fig.width=2, fig.height=2,echo=FALSE,fig.cap="Bacterial Respiration and Production Vial"}
library(png)
library(grid)
img <- readPNG("../figures/C14Tube.png")
 grid.raster(img)
```

# Optimizations

As we start there are a few things we need to optimize.
We are going to start by determining the efficiency of our apparatus. 
We are doing this using 14C-Bicarbonate.
But we need to check a few things first:

1. How much Bicarbonate should we add
2. What is the efficiency of the trap

## 1) Bicarbonate

I did a simple experiment where I added known volumes of the bicarbonate stock to 1500 $\mu$L of scintillation cocktail (1, 5, and 10 $\mu$L).
Based on this I discovered that the LSC could only measure 1 $\mu$L max.
The CPM for this voucher was ~2,300,000 cpm.

## 2) Trap Efficiency

```{r, results='hide'}
# Import Data
trap.data <- read.csv("../notes/20150921_BRBP.csv", header=T)

# Raw Data For Intra Vial Test
in1  <- trap.data[trap.data$Sample == "Tube1_in" & trap.data$Time == "1", 3]
out1 <- trap.data[trap.data$Sample == "Tube1_out" & trap.data$Time == "1", 3]

# Define Theoretical Value
voucher <- trap.data[trap.data$Sample == "Voucher" & trap.data$Time == "1", 3]

# Correct For Vol
in1.c <- in1 * 5 # used 100ul of 500ul
out1.c <- out1 * 50 # used 100ul of 5000ul

# Percent of Theoretical
in1.p <- in1.c / voucher
out1.p <- out1.c / voucher

# Calculate mean and sem
means <- c("Trap" = mean(in1.p), "Residual" = mean(out1.p))
ses <- c("Trap" = se(in1.p), "Residual" = se(out1.p))

# Plot
png(filename="../figures/Recovery1.png",
    width = 800, height = 800, res = 96*2)

par(mar=c(3,5,0.5,0.5), oma=c(1,1,1,1)+0.1, lwd=2)
bp_plot <- barplot(means, ylab = "Percent Recovery", 
                   ylim = c(0, 1), lwd=3, yaxt="n", col="gray",
                   cex.lab=1.5, cex.names = 1.25)
arrows(x0 = bp_plot, y0 = means, y1 = means - ses, angle = 90,
       length=0.1, lwd = 2)
arrows(x0 = bp_plot, y0 = means, y1 = means + ses, angle = 90,
       length=0.1, lwd = 2)
axis(side = 2, labels=T, lwd.ticks=2, las=2, lwd=2)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="14C-Bicarbonate Recovery"}
library(png)
library(grid)
img <- readPNG("../figures/Recovery1.png")
 grid.raster(img)
```

